---
title: "Data Visualization in R"
subtitle: "Barplot with Error Bars"
author: ""
format: 
  html:
    embed-resources: true
    code-overflow: wrap
execute:
  error: true
---

```{r}
library(tidyverse)

library(janitor)

library(readxl)

library(palmerpenguins)
```

### Import Data & Format Column Names

```{r}
tortoise <- read_csv("https://raw.githubusercontent.com/ckaoxvii/CIS-351-2026-Spring/refs/heads/main/tortoise.csv")

tortoise <- tortoise |> 
  clean_names() |> 
  rename(
    tortoise_id = tortoise,
    latitude_n = lattitude_n
  )

names(tortoise)
```

```{r}
tortoise_count <- tibble(
  site = c('Dune', 'Peninsula'),
  'Juvenile' = c(4, 4),
  'Subadult' = c(6, 5),
  'Adult' = c(5, 5)
)

tortoise_count
```

### Barplot with Error Bars

Barplot with error bars provides how precise a measurement is for an aggregated variable such as the mean value of a variable with distinct subgroups.

:::{.callout-note icon='false'}
## Example 1.

Plot the mean body mass of the penguins by species and add the error bars to each bar.

_Solution._

```{r}
penguins_bm_sum <- penguins |> 
  summarize(
  )

penguins_bm_sum
```

```{r}
plot_bm_bar <- penguins_bm_sum |> 
  ggplot() +
  geom_bar(
  ) +
  geom_errorbar(
  )

plot_bm_bar
```
:::

:::{.callout-tip icon='false'}
## Exercise 1.

_Solution._

```{r}

```
:::

### Scatter Plots

:::{.callout-note icon='false'}
## Example 2.

Make a scatter plot of the penguins' body mass vs flipper length.

_Solution._

**Method 1.** Base R.

```{r}
plot_bm_fl <- plot(
)

plot_bm_fl

bm_fl_model <- lm()

abline()
```

**Method 2.** `{ggplot2}`

```{r}
plot_bm_fl_2 <- penguins |> 
  ggplot(aes()) +
  geom_point() +
  geom_smooth()

plot_bm_fl_2
```
:::

:::{.callout-tip icon='false'}
## Repeat Example 2 with penguins' bill length and bill depth.

_Solutions._

**Method 1.** Base R.

```{r}

```

**Method 2.** `{ggplot2}`
:::