---
title: "`filter_out()` and the Pipe Operator"
subtitle: '03/11/2026'
author: ''
format: 
  html:
    embed-resources: true
    code-overflow: wrap
execute:
  error: true
---

```{r}
#| message: false

library(tidyverse)

library(janitor)

library(readxl)

library(palmerpenguins)
```

### Import Data

```{r}
tortoise <- read_csv("https://raw.githubusercontent.com/ckaoxvii/CIS-351-2026-Spring/refs/heads/main/tortoise.csv")

patients <- tibble(
  name = c("Anne", "Mark", "Sarah", "Davis", "Max", "Derek", "Tina"),
  deceased = c(FALSE, TRUE, NA, TRUE, NA, FALSE, TRUE),
  date = c(2005, 2010, NA, 2020, 2010, NA, NA)
)
```

### Format Column Names

```{r}
```

### The `filter_out()` Function

The `filter_out()` function is the complement of the `filter()`, where `filter_out()` drops the rows that satisfy certain logical expressions instead of keeping them.

:::{.callout-note icon='false'}
## Example 1.

Use `filter_out()` to remove the patients who were deceased before 2012.

```{r}
```
:::

:::{.callout-tip icon='false'}
## Exercise 1.

Use `filter_out()` to remove juvenile tortoises whose mass is below the mean mass of all tortoises.

```{r}
```
:::

### The Pipe Operator: `|>`

The pipe operator `|>` lets us chain multiple functions or operations in sequence without using function composition notation.

:::{.callout=note icon="false"}
## Example 2.

Select the `site`, `tortoise`, `class`, and `ticks_count` columns from `tortoise`. Then rename the `tortoise` column to `tortoise_id` and remove adult tortoises with a tick count greater than 5.

_Method 1:_ Without using `|>`
```{r}
```
:::