---
title: "`filter()`"
subtitle: '03/04/2026'
author: ''
format: 
  html:
    embed-resources: true
---

```{r}
#| message: false

library(tidyverse)

library(janitor)

library(readxl)

library(palmerpenguins)
```

### Import Data

```{r}
tortoise <- read_csv("https://raw.githubusercontent.com/ckaoxvii/CIS-351-2026-Spring/refs/heads/main/tortoise.csv")
```

### Format Column Names

```{r}
names(___)

tortoise <- ___(tortoise)

names(___)
```

:::{.callout-tip icon='false'}
## Exercise 0.

Examine the column names of the `tortoise` data frame. Do you notice any typographical errors?

_**Solution.**_

`lattitude_n` is misspelled. We can use the `rename()` function to correct this error.

```{r}
tortoise <- rename(___, ___)

names(___)
```
:::

### Basic Data Exploration

```{r}
str(___)
```
:::{.callout-caution icon='false'}
## `unique()`

The `unique()` function is useful for listing distinct values in a vector or a column of a data frame.
:::

:::{.callout-note icon='false'}
## Example 1.

Use `unique()` to determine the distinct values in the `site` column of `tortoise`.

_**Solution.**_

```{r}
unique(___)
```
:::

:::{.callout-tip icon='false'}
## Exercise 1.

Find the unique values in the remaining columns of `tortoise`.

_**Solution.**_

```{r}

```
:::

:::{.callout-note icon='false'}
## Example 2.

Calculate the descriptive statistics for `mass_g` and summary of `class`.

```{r}
summary(___)
```

```{r}
___(tortoise$class)
```
:::

:::{.callout-tip icon='false'}
## Exercise 2.

Calculate the descriptive statistics of the remaining numeric variables and summarize character variables in `tortoise`.

```{r}

```
:::

### `filter()`

:::{.callout-caution icon='false'}
## `filter()`

The `filter()` function keeps the rows of a data frame that satisfy certain logical conditions. The syntax is `filter(.data, ...)`, where `.data` is a data frame and `...` logical expressions.
:::

:::{.callout-note icon='false'}
## Example 3.

a. Subset out the tortoises from the Dunes site.
b. Subset out the tortoises whose mass is greater than 500 g.
c. Subset out all the adult male tortoises.

_**Solutions.**_

a. 
```{r}
tortoise_dunes <- filter(___)

___(tortoise_dunes$site) # check that Dunes is the only value in `site` column.
```

b.
```{r}
tortoise_heavy <- filter(___)

___(tortoise_heavy$mass_g)
```

c.
```{r}
tortoise_adult_male <- filter(___)

___jk
```
:::

:::{.callout-tip icon='false'}
## Exercise 3.

a. Subset out the tortoises from Peninsula site.
b. Subset out the tortoises whose straight carapace length is less than 150 mm or greater than 200 mm.
c. Subset out the juvenile tortoises with a tick count greater than 5.
d. Create a data frame that contains only the non-juvenile tortoises.


_**Solutions.**_

a.
```{r}

```

b.
```{r}

```

c.
```{r}

```

d.
```{r}

```
:::

:::{.callout-note icon='false'}
## Exercise 4.

Filter out rows where the patient is deceased and the year was before 2012.

```{r}
patients <- tibble(
  name = c("Anne", "Mark", "Sarah", "Davis", "Max", "Derek", "Tina"),
  deceased = c(FALSE, TRUE, NA, TRUE, NA, FALSE, TRUE),
  date = c(2005, 2010, NA, 2020, 2010, NA, NA)
)

patients
```

_**Solution.**_ 

```{r}

```
:::