---
title: "Data Visualization in R"
subtitle: "Histogram and Bar Plot"
date: "04/01/2026"
author: ""
format: 
  html:
    embed-resources: true
    code-overflow: wrap
execute:
  error: true
---

```{r}
library(tidyverse)

library(janitor)

library(readxl)

library(palmerpenguins)
```

### Import Data & Format Column Names

```{r}
tortoise <- read_csv("https://raw.githubusercontent.com/ckaoxvii/CIS-351-2026-Spring/refs/heads/main/tortoise.csv")

tortoise <- tortoise |> 
  clean_names() |> 
  rename(
    tortoise_id = tortoise,
    latitude_n = lattitude_n
  )

names(tortoise)
```

```{r}
tortoise_count <- tibble(
  site = c('Dune', 'Peninsula'),
  'Juvenile' = c(4, 4),
  'Subadult' = c(6, 5),
  'Adult' = c(5, 5)
)

tortoise_count
```

```{r}
datasaurus_dozen <- read_csv("https://raw.githubusercontent.com/ckaoxvii/CIS-351-2026-Spring/refs/heads/main/datasaurus_dozen.csv") |> 
  rename(shape = dataset)
```

### Why Data Visualization?

Run the following code chunk. What can you say about the data by `shape`?

```{r}
names(datasaurus_dozen)

datasaurus_sum <- datasaurus_dozen |> 
  summarize(
    x_mean = round(mean(x, na.rm = TRUE), 2),
    y_mean = round(mean(y, na.rm = TRUE), 2),
    x_sd = round(sd(x, na.rm = TRUE), 2),
    y_sd = round(sd(y, na.rm = TRUE), 2),
    .by = shape
  )

datasaurus_sum
```

```{r}
datasaurus_dozen |> 
  ggplot(aes(x = x, y = y, color = shape))+
    geom_point() +
    theme_void() +
    theme(legend.position = "none")+
    facet_wrap(~shape, ncol = 3)
```

### Histograms

#### Histograms in Base R

:::{.callout-note icon='false'}
## Example 1.

Plot the distribution of the penguins' body mass in base R.

```{r}
```
:::

:::{.callout-tip icon='false'}
## Exercise 1.

Plot the distribution of the penguins' flipper length in base R.

```{r}
```
:::

#### Histogram in `{ggplot2}`

:::{.callout-note icon='false'}
## Example 2.

Plot the distribution of the penguins' body mass using `{ggplot2}`.

```{r}
```
:::

:::{.callout-tip icon='false'}
## Exercise 2.

Plot the distribution of the penguins' flipper length using `{ggplot2}`.

```{r}
```
:::

### Bar Plots

#### Bar Plots in Base R

:::{.callout-note icon='false'}
## Example 3.

Create a bar plot in base R that shows the count of penguins by species.

```{r}
```
:::


:::{.callout-note icon='false'}
## Exercise 3.

Create a bar plot in base R that shows the count of penguins by island.

```{r}
```
:::

#### Bar Plots Using `{ggplot2}`

:::{.callout-note icon='false'}
## Example 4.

Create a bar plot using `{ggplot2}` that shows the count of penguins by species.

```{r}
```
:::

:::{.callout-note icon='false'}
## Exercise 4.

Create a bar plot using `{ggplot2}` that shows the count of penguins by island.

```{r}
```
:::