---
title: "The `mutate()` Function and Additional Helper Functions"
subtitle: '03/23/2026'
author: ''
format: 
  html:
    embed-resources: true
    code-overflow: wrap
execute:
  error: true
---

```{r}
#| message: false

library(tidyverse)

library(janitor)

library(readxl)

library(palmerpenguins)
```

### Import Data & Format Column Names

```{r}
tortoise <- read_csv("https://raw.githubusercontent.com/ckaoxvii/CIS-351-2026-Spring/refs/heads/main/tortoise.csv")

tortoise <- tortoise |> 
  clean_names() |> 
  rename(tortoise_id = tortoise)

names(tortoise)
```

```{r}
patients <- tibble(
  name = c("Anne", "Mark", "Sarah", "Davis", "Max", "Derek", "Tina"),
  deceased = c(FALSE, TRUE, NA, TRUE, NA, FALSE, TRUE),
  date = c(2005, 2010, NA, 2020, 2010, NA, NA)
)
```

### Optional Arguments of the `mutate()` Function

Recall the syntax for `mutate()` is as follows

```r
mutate(
  .data,
  ...,
  .by = NULL,
  .keep = c("all", "used", "unused", "none"),
  .before = NULL,
  .after = NULL
)
```

In the previous class, we saw that the `.by` argument can be used to add or modify variables by group. We will explore the remaining optional arguments of `mutate()`.

#### The `.keep` Option

This option lets us control which columns we want to retain in the output.

  * `"all"` retains all columns from `.data`; this is the default.
  * `"used"` retains only the columns used in the `...` step(s) of `mutate()`. This is useful for checking your work.
  * `"unused"` retains only the columns _not_ used in the `...` step(s) of `mutate()`. This is useful if new columns are generated, but the old columns used to generate the new ones are no longer needed.
  * `"none"` retains only the new columns generated in the `...` step(s) of `mutate()`.

:::{.callout-note icon="false"}
## Example 1. 

Use `mutate()` to create a new variable called `straight_carapace_l_cm`, then explore each of `.keep` options.

```{r}
tortoise_keep_all <- 

names(tortoise_keep_all)
```

```{r}
tortoise_keep_used <- 

names(tortoise_keep_used)
```

```{r}
tortoise_keep_unused <- 

names(tortoise_keep_unused)
```

```{r}
tortoise_keep_none <- 

names(tortoise_keep_none)
```
:::

#### The `.before` and `.after` Options

These two optional arguments let us control the location of the new or modified column(s).

:::{.callout-note icon='false'}
## Example 2.

Create the variable `carapace_lw_ratio` by calculating the ratio of `straight_carapace_length_mm` to `carapace_width_mm`. Then reposition `carapace_lw_ratio` to be after `carapace_width_mm`.

_Solution._ Note that there are two ways of performing the relocation of `carapace_lw_ratio`, we can position it after `carapace_width_mm`, or before `height_mm`.

_Method 1:_ Using '.after'
```{r}
tortoise_2a <- 

names(tortoise_2a)
```

_Method 2:_ Using `.before`

```{r}
tortoise_2b <- 

names(tortoise_2b)
```

```{r}
names(tortoise_2a) == names(tortoise_2b)
```
:::

## Useful Helper Functions

The following are some useful helper functions from the `{dplyr}` and `{tidyselect}` packages that expand the capabilities of `filter()` and `mutate()`.

* `across()`: Apply the same transformation to multiple columns.
* `if_any()`: Select any columns when the predicate is `TRUE`.
* `if_all()`: Select _all_ columns when the predicate is `TRUE`.
* `starts_with()`: Starts with an exact prefix.
* `ends_with()`: Ends with an exact suffix.
* `contains()`: Contains a literal string.
* `matches()`: Matches a regular expression.
* `num_range()`: Matches a numerical range.

:::{.callout-note icon='false'}
In the `penguins` data frame, convert all columns measured in mm to meters.

_Solution._
```{r}
penguins_meters <- 

head(penguins_meters)
```
:::