---
title: "Data Visualization in R"
subtitle: "Scatter Plots & Line Graphs"
author: ""
format: 
  html:
    embed-resources: true
    code-overflow: wrap
execute:
  error: true
---

```{r}
library(tidyverse)

library(janitor)

library(readxl)

library(palmerpenguins)

library(ggpubr)
```

### Import Data & Format Column Names

```{r}
tortoise <- read_csv("https://raw.githubusercontent.com/ckaoxvii/CIS-351-2026-Spring/refs/heads/main/tortoise.csv")

tortoise <- tortoise |> 
  clean_names() |> 
  rename(
    tortoise_id = tortoise,
    latitude_n = lattitude_n
  )

names(tortoise)
```

```{r}
tortoise_count <- tibble(
  site = c('Dune', 'Peninsula'),
  'Juvenile' = c(4, 4),
  'Subadult' = c(6, 5),
  'Adult' = c(5, 5)
)

tortoise_count
```

## Adding Values to Scatter Plots

:::{.callout-note icon='false'}
## Example 1 

Add the correlation coefficient, the regression-line equation, and the $p$-value to the scatter plot of penguins' body mass versus flipper length.

_Solution._ We will use the `ggscatter()` function from `{ggpubr}` package.

```{r}
```
:::

:::{.callout-tip icon='false'}
## Exercise 1.

Repeat Example 1 with a scatter plot of penguins' bill depth and bill length.

```{r}

```
:::

## Box Plots

A box plot is a graphical method for visualizing the distribution of a numeric variable across different groups.

:::{.callout-note icon='false'}
## Example 2.

a. Create a box plot of penguins' body mass by species.
b. Create a box plot of penguins' body mass by species where the box is colored by species.

_Solutions._

a. 
```{r}
```

b. 
```{r}
```
:::

:::{.callout-tip icon='false'}
## Exercise 2. 

Repeat Example 2 for penguins' flipper length.
:::

### Grouped Box Plots

:::{.callout-note icon='false'}
## Example 3.

Create a grouped box plot for the penguins' body mass by species and grouped by sex.

```{r}
```

```{r}
```

```{r}
```
:::

:::{.callout-tip icon='false'}
## Exercise 3.

Repeat Example 3 with penguins' flipper length.
:::
