---
title: "Data Visualization in R"
subtitle: "Stacked Bar Plots and Customizing Plots"
author: ""
format: 
  html:
    embed-resources: true
    code-overflow: wrap
execute:
  error: true
---

```{r}
library(tidyverse)

library(janitor)

library(readxl)

library(palmerpenguins)

library(ggpubr)
```

### Import Data & Format Column Names

```{r}
tortoise <- read_csv("https://raw.githubusercontent.com/ckaoxvii/CIS-351-2026-Spring/refs/heads/main/tortoise.csv")

tortoise <- tortoise |> 
  clean_names() |> 
  rename(
    tortoise_id = tortoise,
    latitude_n = lattitude_n
  )

names(tortoise)
```

```{r}
tortoise_count <- tibble(
  site = c('Dune', 'Peninsula'),
  'Juvenile' = c(4, 4),
  'Subadult' = c(6, 5),
  'Adult' = c(5, 5)
)

tortoise_count
```

## Stacked Bar Plot

:::{.callout-note icon='false'}
## Example 1.

Create a stacked bar plot of penguins' sex by species.

_Solution._

```{r}
```
:::

:::{.callout-note icon='false'}
## Example 2.

Create a stacked bar plot of penguins' species by island.

_Solution._

```{r}
```
:::

### Customize Plots

:::{.callout-note icon='false'}
## Example 3.

Create a scatter plot of penguins' body mass vs flipper length. Change the background theme and axis labels.

```{r}
```
:::

:::{.callout-note icon='false'}
## Example 4.

Repeat Example 1, but change the color of the bars.

```{r}
```
:::