---
title: "The `summarize()` Function"
subtitle: '03/25/2026'
author: ''
format: 
  html:
    embed-resources: true
    code-overflow: wrap
execute:
  error: true
---

```{r}
#| message: false

library(tidyverse)

library(janitor)

library(readxl)

library(palmerpenguins)

library(tinytable)
```

### Import Data & Format Column Names

```{r}
tortoise <- read_csv("https://raw.githubusercontent.com/ckaoxvii/CIS-351-2026-Spring/refs/heads/main/tortoise.csv")

tortoise <- tortoise |> 
  clean_names() |> 
  rename(
    tortoise_id = tortoise,
    latitude_n = lattitude_n
  )

names(tortoise)
```

```{r}
patients <- tibble(
  name = c("Anne", "Mark", "Sarah", "Davis", "Max", "Derek", "Tina"),
  deceased = c(FALSE, TRUE, NA, TRUE, NA, FALSE, TRUE),
  date = c(2005, 2010, NA, 2020, 2010, NA, NA)
)
```

## Refactoring Data

:::{.callout-note icon='false'}
## Example 0.

Refactor the penguins data frame to ensure that the values are consistent.

```{r}

```
:::

## Summarizing Data

`summarize()` (or `summarise()`, if you prefer the King's English) is used to create a new data frame containing summary values of your choice. The syntax for `summarize()` is

```r
summarize(
  .data,
  ...,
  .by = NULL
)
```

:::{.callout-note icon='false'}
## Example 1.

Using the `penguins` data set, use `summarize()` to calculate the following:

a. the mean flipper length of all penguins.
b. the mean flipper length of all penguins, by species.
c. the mean flipper length of all penguins, by species and sex.

_Solutions._

a. The mean flipper length of all penguins.

```{r}
```

b. The mean flipper length of all penguins, by species.

```{r}
```

c. The mean flipper length of all penguins, by species and sex.

```{r}
```
:::

:::{.callout-tip icon='false'}
## Exercise 1.

Using the `tortoise` data set, calculate the following:

a. The mean and standard deviation of tortoise mass.
b. The mean and standard deviation of tortoise mass, by site.
c. The mean and standard deviation of tortoise mass, by site and class.

_Solutions._

a.
```{r}
```

b.
```{r}
```

c.
```{r}
```
:::

## The `{tinytable}` Package

The `{tinytable}` package provides functions to format data frame output.

:::{.callout-note icon='false'}
## Example 2.

Format the tables from Example 1 using `tt()` from `{tinytable}`.

```{r}
```

```{r}
```

```{r}
```
:::

:::{.callout-tip icon='false'}
## Exercise 2. 

Repeat Example 2 using the summary tables from Exercise 1.

```{r}
```

```{r}
```

```{r}
```
:::