---
title: "Statistical Tests in R"
subtitle: "$t$-Test and ANOVA"
author: ""
format: 
  html:
    embed-resources: true
    code-overflow: wrap
execute:
  error: true
---

```{r}
library(tidyverse)

library(janitor)

library(readxl)

library(palmerpenguins)

library(ggpubr)
```

## One-Sample $t$-Test

A one-sample $t$-test is a hypothesis test used to determine whether a sample mean differs from a population mean.

:::{.callout-note icon='false'}
## Example 1.

Conduct a one-sample $t$-test to determine whether the mean body mass of the chinstrap penguins is the same as the mean body mass of all penguins. Use $\alpha = 0.5$ as the significance level.

_Solution._

First we need to know what the mean body mass of all penguins:

```{r}
```

Our hypotheses are

* $H_0: $
* $H_a: $

```{r}
```

:::

By default, the `t.test` function uses a two-sided alternative hypothesis. We can change this using the optional argument `alternative = "less"` or `alternative = "greater"`.

:::{.callout-note icon='false'}
## Example 2.

Conduct a two-sample $t$-test to determine whether the gentoo penguins' mean body mass is greater than the mean body mass of all penguins. Use $\alpha = 0.1$.

_Solution._

The hypotheses are:

* $H_0: $
* $H_a: $

```{r}
```

:::

## Two-Sample $t$-Test

A two-sample $t$-test is a hypothesis test used to determine whether the means of two independent populations are significantly different based on sample data. 

:::{.callout-note icon='false'}
## Example 3.

Conduct a two-sample $t$-test to determine whether the mean body mass of chinstrap and adelie penguins are different.

_Solution._

Let $\mu_c$ and $\mu_a$ to denote the mean body mass of the chinstrap and adelie penguins, respectively. The hypotheses are

* $H_0: $
* $H_a: $

```{r}
```

```{r}
```
:::