---
title: "Statistical Tests in R"
subtitle: "Tests for Categorial Data"
author: ""
format: 
  html:
    embed-resources: true
    code-overflow: wrap
execute:
  error: true
---

```{r}
#| message: false
 
library(tidyverse)

library(janitor)

library(readxl)

library(palmerpenguins)

library(ggpubr)
```

## Chi-Square Test of Independence

The chi-square test of independence is used to determine if there is a significant association between two categorical variables. It compares the observed frequencies in each category to the frequencies we would expect if there were no association between the variables.

The hypotheses for the chi-square test of independence are:

* $H_0$: There is no association between the two categorical variables (knowing the value of one variable does not provide any information about the other variable).
* $H_a$: There is an association between the two categorical variables (knowing the value of one variable provides information about the other variable).

:::{.callout-note icon="false"}
## Example 1.

Use the chi-square test of independence to determine if there is an association between penguins' species and their body mass. 

_Solution._

The hypotheses for this test are:

* $H_0$: There is no association between penguins' species and their body mass
* $H_a$: There is an association between penguins' species and their body mass

The body mass variable is a continuous variable, so we need to categorize it before we can perform the chi-square test of independence. We will use the following thresholds to categorize body mass:

* Light: if the body mass is in the bottom 25% of the data
* Medium: if the body mass is in the middle 50% of the data
* Heavy: if the body mass is in the top 25% of the data

**Step 1.** Create a new variable `body_mass_category` in the `penguins` dataset that categorizes body mass into "Light", "Medium", and "Heavy" based on the thresholds defined above.
```{r}
```

**Step 2.** Create a contingency table of the `species` and `body_mass_category` variables. There are mutliple ways of doing this, the easiest one is to use the (base R) `table()` function.

```{r}
```

**Step 3.** Create a stacked bar plot to visualize the relationship between `species` and `body_mass_category`.

```{r}
```

You can also use an unstacked bar plot to visualize the same relationship between `species` and `body_mass_category`.

```{r}
```

**Step 4.** Perform the chi-square test of independence using the `chisq.test()` function.

```{r}
```

**Step 5.** Interpret the results of the chi-square test of independence.


:::

## Goodness-of-Fit Test

The goodness-of-fit test is used to determine if a sample data matches a population with a specific distribution. It compares the observed frequencies in each category to the frequencies we would expect if the data followed the specified distribution.

The hypotheses for the goodness-of-fit test are:
* $H_0$: The observed data follows the specified distribution.
* $H_a$: The observed data does not follow the specified distribution.

:::{.callout-note icon="false"}
## Example 2.

Use the goodness-of-fit test to determine if the distribution of penguins' species in the `penguins` dataset follows a uniform distribution (i.e., each species is equally represented).

_Solution._

The hypotheses for this test are:

* $H_0$: The distribution of penguins' species follows a uniform distribution.
* $H_a$: The distribution of penguins' species does not follow a uniform distribution.

**Step 1.** Create a table of the `species` variable to get the observed frequencies of each species.

```{r}
```

**Step 2.** Create a bar plot to visualize the distribution of penguins' species.

```{r}
```

**Step 3.** Perform the goodness-of-fit test using the `chisq.test()` function. Since we are testing against a uniform distribution, we need to specify the expected probabilities for each category (in this case, each species should have an equal probability of 1/3).

```{r}
```

**Step 4.** Interpret the results of the goodness-of-fit test.


:::
